---
title: "Version Control with Git"
#author: "Liming Wang"
#date: "8/21/2017"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Why learn version control?
- Allows for collaborators to work in parallel.
- Eliminates the need to come up with file names to indicate a "final" versus a "final-final" file version.

## Install Git {#install-git}
For installation directions please refer to the [Setup, overview, and introduction](01-part1-overview.html) page. 

## Lessons
- [A quick overview of Version Control](http://swcarpentry.github.io/git-novice/01-basics/index.html)
- Start off by registering a github account at https://github.com
- Use your github user name and email to config git. Open a Shell/Command line window from RStudio menu **Tools/Shell...**, and modify and run the following commands:
```{r, eval=FALSE}
git config --global user.name "Liming Wang"
git config --global user.email "lmwang+datascience@gmail.com"
git config --global color.ui "auto"
```

- [Creating a repository](http://swcarpentry.github.io/git-novice/03-create/index.html)
- [Tracking changes](http://swcarpentry.github.io/git-novice/04-changes/index.html)
- [Exploring history](http://swcarpentry.github.io/git-novice/05-history/index.html)
- [Ignoring things](http://swcarpentry.github.io/git-novice/06-ignore/index.html)
- [Remotes in GitHub](http://swcarpentry.github.io/git-novice/07-github/index.html)
- [Collaborating](http://swcarpentry.github.io/git-novice/08-collab/)
- [Conflicts](http://swcarpentry.github.io/git-novice/09-conflict/)
- [Review of common git commands](http://happygitwithr.com/usage-git-cmds.html)
- [Hosting](http://swcarpentry.github.io/git-novice/13-hosting/index.html)

## Optional lessons and additional resources
- [Open Science](http://swcarpentry.github.io/git-novice/10-open/index.html)
- [Licensing](http://swcarpentry.github.io/git-novice/11-licensing/index.html)
- [Citation](http://swcarpentry.github.io/git-novice/12-citation/index.html)
- [Cache credentials](http://happygitwithr.com/credential-caching.html)

## Resources
1. [Happy Git with R](http://happygitwithr.com/)
1. [try Git](https://try.github.io)
1. [Pro Git](https://git-scm.com/book/en/v2), a book by Scott Chacon and Ben Straub
