---
title: "Data manipulation with dplyr"
output: 
  html_document:
    toc: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The tidyverse packages
- Website: http://www.tidyverse.org/packages/
- Install with `install.packages("tidyverse")`

Tidyverse is a package of packages that shares a common philosophy designed for data wrangling, exploring, and visualization. Of the many packages associated with tidyverse we've already covered a brief introduction to `ggplot2`, and will next cover `dplyr` and `tidyr`.

dplyr is a grammar of data manipulation, providing a consistent set of verbs that help you solve the most common data manipulation challenges:

- `mutate()` adds new variables that are functions of existing variables
- `select()` picks variables based on their names.
- `filter()` picks cases based on their values.
- `summarise()` reduces multiple values down to a single summary.
- `arrange()` changes the ordering of the rows.

These all combine naturally with `group_by()` which allows you to perform any operation "by group". You can learn more about them in `vignette("dplyr")`. As well as these single-table verbs, dplyr also provides a variety of two-table verbs, which you can learn about in `vignette("two-table")`.

## Pipe operator
`%>%` pipes an object forward into a function or call expression

- Basic piping
    - x %>% f is equivalent to f(x)
    - x %>% f(y) is equivalent to f(x, y)
    - x %>% f %>% g %>% h is equivalent to h(g(f(x)))

- Using `%>%` with unary function calls

```{r, echo=TRUE}
require(tidyverse)
iris %>% head
iris %>% tail
```

More information available at: http://magrittr.tidyverse.org/

RStudio keyboard shortcut for `%>%`:

- Ctrl-Shift-M (Windows)
- Command-Shift-M (Mac)

## Lesson
1. [Dataframe manipulation with dplyr](http://swcarpentry.github.io/r-novice-gapminder/13-dplyr/index.html)

## Optional lesson
1. [Programming with dplyr](http://dplyr.tidyverse.org/articles/programming.html)

## Resources
1. [Data wrangling cheat sheet](https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf)
1. [Relational data](https://r4ds.had.co.nz/relational-data.html)
