---
title: "Formatting dates"
output: 
  html_document:
    toc: TRUE
---

Working with time-series data can be a challenge for new and experienced R users. You will often have to format the date, time, and timezone when working with raw data. R does not automatically recognize date-time formats and there are many formats for representing date-time (e.g. yyyy-mm-dd, mm-dd-yy, mm/dd/yyyy hh:mm:ss).

`lubridate` is a handy package that is installed as part of the `tidyverse` installation but does not automatically load when you call for the `tidyverse` package (`library(tidyverse)`). You have to explicitly call the package when you need it with `library(tidyverse)`.

## Lesson
Download the sample data set from [here](https://raw.githubusercontent.com/cities/datascience2019/master/data/leesferry.csv) (left-click & Save link as...), and move it to the data folder of your working drive.

Load the library and sample dataset.
```{r setup, results="hide", message=FALSE}
library(lubridate)
library(dplyr)

leesferry <- read.csv("data/leesferry.csv", stringsAsFactors = F)
```

Check the structure of your dataframe and you'll notice the `date.time` attribute is a character string. You'll also notice that the timezone is MST.
```{r, echo = TRUE}
str(leesferry)
```

When assigning timezones to your date-time attribute use the Olson name which can be called by `OlsonNames()` of which there are over 600. You can also look at a list [here](https://en.wikipedia.org/wiki/List_of_tz_database_time_zones).

To format the `date.time` attribute with the appropriate timezone use the following `lubridate` function:
```{r ymd_hms_tz, results = "hide"}
leesferry$date.time <- ymd_hms(leesferry$date.time, tz = "US/Mountain")
```

To extract the year, month, or day of a date:
```{r extract from date, results = "hide"}
# Extract and create new column for year
leesferry$year <- year(leesferry$date.time)

# Extract and create new columns for month and day of week
leesferry_dow <- leesferry %>%
  mutate(month = month(date.time)) %>%
  mutate(dow = wday(date.time, label = T, abbr = T))
```

There are a number of functions with lubridate that can make working with time-series data including but not limited to:

- rounding date-times (helpful with aggregations)
- calculating intervals, periods, and durations

Check out the [vignette](https://cran.r-project.org/web/packages/lubridate/lubridate.pdf) for these additional functions.

## Resources
1. [Dates and times](https://r4ds.had.co.nz/dates-and-times.html)
1. [Lubridate cheat sheet](https://github.com/rstudio/cheatsheets/raw/master/lubridate.pdf)