---
title: "Formatting dates"
output: 
  html_document:
    toc: TRUE
---

Working with time-series data can be a challenge for new and experienced R users. You will often have to format the date, time, and timezone when working with raw data. R does not automatically recognize date-time formats and there are many formats for representing date-time (e.g. yyyy-mm-dd, mm-dd-yy, mm/dd/yyyy hh:mm:ss).

`lubridate` is a handy package that is installed as part of the `tidyverse` installation but does not automatically load when you call for the `tidyverse` package (`library(tidyverse)`). You have to explicitly call the package when you need it.

## Lesson

Use one of the Biketown data files you downloaded the other day in the writing functions exercise. Or use `source` to use the function to get Biketown data for 06/2018 through 08/2018.
```{r setup}
library(lubridate)
library(dplyr)
library(ggplot2)

source("/home/tammy/Documents/ds19-class/code/biketown-example.R")
get_data(start = "06/2018", end = "08/2018")

# Function to read in all files and combine into one dataframe
# This function only works if you explicitly set the working drive to
# where the data is being stored

# Aftering running the function, make sure to set the working drive
# back to the folder where your .Rproj file is stored.

setwd("/home/tammy/Documents/ds19-class/data/biketown")
folder <- "/home/tammy/Documents/ds19-class/data/biketown"
filenames <- list.files(path = folder, pattern = "*.csv", all.files = FALSE, full.names = FALSE,
                        recursive = FALSE, ignore.case = FALSE)
read_csv_filename <- function(filenames){
  ret <- read.csv(filenames, stringsAsFactors = F, 
                  strip.white = T, na.strings = "")
  ret$Source <- filenames
  ret
}
bike_raw <- plyr::ldply(filenames, read_csv_filename)

setwd("/home/tammy/TREC/datascience2019/")
```

```{r data_wrangling}
# check data structure
str(bike_raw)

# create new columns `start.datetime` and `end.datetime`
bike_df1 <- bike_raw %>%
  mutate(start.datetime = paste(StartDate, StartTime, sep = " "),
         end.datetime = paste(EndDate, EndTime, sep = " "))

# convert `start.datetime` and `end.datetime` into date time format with appropriate timezone
bike_df1$start.datetime <- mdy_hm(bike_df1$start.datetime, tz = "America/Los_Angeles")
bike_df1$end.datetime <- mdy_hm(bike_df1$end.datetime, tz = "America/Los_Angeles")

# convert `Duration` into a useable format
bike_df1$Duration <- hms(bike_df1$Duration)
# this throws a warning about NA's

# checking for NAs in `bike_raw$Duration`
sum(is.na(bike_raw$Duration))
```

There are three functions in `lubridate` that seem synonomous but define very different actions:

1. duration: span of time measured in seconds, and there is no start date involved (see above example useage)
1. interval: measures between two specific time points (in seconds)
1. period: measures time span in units larger than seconds, handy for when accounting for daylight saving times, leap years

```{r datetime}
# calculate interval
bike_df1$interval <- interval(bike_df1$start.datetime, bike_df1$end.datetime)
head(bike_df1$interval)

# calculate duration
bike_df1$duration.all <- as.duration(bike_df1$interval)
head(bike_df1$duration.all)

# calculate period
bike_df1$period <- as.period(bike_df1$duration.all)
head(bike_df1$period)
```

```{r aggregate_datetime}
# using floor_date to help aggregate data
# want weekly mean distance traveled

bike_wkagg <- bike_df1 %>%
  mutate(week.datetime = floor_date(start.datetime, unit = "week")) %>%
  group_by(week.datetime) %>%
  summarise(weekly.meandist = mean(Distance_Miles))
str(bike_wkagg)
bike_wkagg$week.datetime <- as.Date(bike_wkagg$week.datetime)

weekly_meandist_fig <- bike_wkagg %>%
  ggplot(aes(x = week.datetime, y = weekly.meandist)) +
  geom_bar(stat = "identity", fill = "orange") +
  scale_x_date(date_breaks = "1 week") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
weekly_meandist_fig
```

There are a few important regular expression (Regex) base R functions that are handy to know when working with strings:

1. `grep` and `grepl`: `grep` looks for a match within a string vector and returns an indice of matches. `grepl` looks for a match within a string and returns a logical vector.
1. `sub` and `gsub`: replaces the *first* exact matching chunk of text within a string vector with a specified replacement. `gsub` replaces *all* exact matching chunks of text with a specified replacement.

```{r regular expressions}
# Create three different station categories for the start and end stations
bike_df2 <- bike_df1 %>%
  mutate(start.station.category = if_else(grepl("Community", StartHub), "Community Station",
                                    if_else(grepl("", StartHub), "Outside Station",
                                      "BIKETOWN Station"))) %>%
  mutate(end.station.category = if_else(grepl("Community", EndHub), "Community Station",
                                    if_else(grepl("", EndHub), "Outside Station",
                                      "BIKETOWN Station")))
table(bike_df2$start.station.category)
table(bike_df2$end.station.category)


```

## Resources
Steve Fick's [Regular Expressions](https://d-rug.github.io/blog/2015/regex.fick)