---
title: "Working with dates and regex"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lubridate)
library(dplyr)
library(ggplot2)
```

## Purpose
Explore BikeTown data looking at the mean miles traveled per week during the summer of 2018 (06/2018 through 09/2018.

```{r pull_data}
source("/home/tammy/Documents/ds19-class/code/biketown-example.R")
get_data(start = "06/2018", end = "09/2018")

setwd("/home/tammy/Documents/ds19-class/data/biketown/")

folder <- "/home/tammy/Documents/ds19-class/data/biketown/"

filenames <- list.files(path = folder, pattern = "*.csv", all.files = FALSE,
                        full.names = FALSE,
                        recursive = FALSE, ignore.case = FALSE)

read_csv_filename <- function(filenames){
  ret <- read.csv(filenames, stringsAsFactors = F, 
                  strip.white = T, na.strings = "")
  ret$Source <- filenames
  ret
}

bike_raw <- plyr::ldply(filenames, read_csv_filename)

setwd("/home/tammy/TREC/datascience2019/")

#### Just call in data one file at a time
file1 <- read.csv("/home/tammy/Documents/ds19-class/data/biketown/2018_06.csv", stringsAsFactors = F)
file2 <- read.csv("/home/tammy/Documents/ds19-class/data/biketown/2018_07.csv", stringsAsFactors = F)

## After calling in all your files use bind_rows()
bike_raw <- bind_rows(file1, file2, ...)
```

```{r data_wrangling}
str(bike_raw)

bike_df1 <- bike_raw %>%
  mutate(start.datetime = paste(StartDate, StartTime, sep = " "),
         end.datetime = paste(EndDate, EndTime, sep = " "))
str(bike_df1)

# covert start.datetime and end.datetime into date-time format,
# make sure to include your timezone

# check acceptable timezone values
OlsonNames()
head(bike_df1$start.datetime)

bike_df1$start.datetime <- mdy_hm(bike_df1$start.datetime, tz = "America/Los_Angeles")
bike_df1$end.datetime <- mdy_hm(bike_df1$end.datetime, tz = "America/Los_Angeles")
str(bike_df1)

# convert duration into a useable format
bike_df1$Duration <- hms(bike_df1$Duration)

# Error thrown after converting to duration indicating NAs,
# Check raw dataframe for NAs in Duration column
sum(is.na(bike_raw$Duration))


```

`lubridate` has three synonomous functions but they define very different actions:

1. duration(): time spans are in units of seconds, and there is no start date invovled.
2. interval(): also measured in units of seconds, and provides the difference between start and end.
3. period(): measures time span in units larger than seconds, this is handy when accounting for daylight saving times, and leap years.

```{r duration_calc}
# Calculate interval
bike_df1$interval <- interval(bike_df1$start.datetime, bike_df1$end.datetime)
head(bike_df1$interval)

# Calculate duration
bike_df1$duration.all <- as.duration(bike_df1$interval)
head(bike_df1$duration.all)

# Calculate period
bike_df1$period <- as.period(bike_df1$duration.all)
head(bike_df1$period)

# floor_date() is helpful for aggregating
bike_wkagg <- bike_df1 %>%
  mutate(week.datetime = floor_date(start.datetime, unit = "week"))
head(bike_wkagg$week.datetime)

# Calculate weekly mean distance traveled
wkmean_dt <- bike_wkagg %>%
  group_by(week.datetime) %>%
  summarise(weekly.meandist = mean(Distance_Miles))
View(wkmean_dt)

# Plot your results

wkmean_dt_fig <- wkmean_dt %>%
  ggplot(aes(x = as.Date(week.datetime), y = weekly.meandist)) +
  geom_bar(stat = "identity", fill = "orange") +
  scale_x_date(date_breaks = "1 week") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
wkmean_dt_fig


```





